---
title: "Star Systems Data Analysis"
author: Daniel Rohrbach
format: html
editor: visual
echo: false
message: false
warning: false
---

```{r}
#| message: false
here::i_am("battletech.Rproj")
library(here)
library(ggplot2)
library(dplyr)
library(tidyr)
library(vroom)
library(stringr)
```

[Link to the Github](https://github.com/dauphine-data-daniel/battletech)

```{r}
data <- vroom(here("data", "List+sytems.csv"), na = c("N/A", "NA", "?"), locale = locale(decimal_mark = ",", grouping_mark = "."))
```

## Q3

```{r}
expected_star_systems <- 1085
expected_variables <- 8

actual_star_systems <- nrow(data)
actual_variables <- ncol(data)

numerical_vars <- c("distance", "star temperature", "relative radius", "MASS", "luminosity")
data_types <- sapply(data[numerical_vars], class)

all_numeric <- function(x) {
  all(sapply(x, function(v) v %in% c("numeric", "integer")))
}

colnames(data)[colnames(data) == "MASS"] <- "mass"

```
The dataset contains `r nrow(data)` star systems and `r ncol(data)` variables. The expected number of star systems is `r expected_star_systems` and the expected number of variables is `r expected_variables`.

`r if (actual_star_systems == expected_star_systems && actual_variables == expected_variables) "The actual counts match the expected counts." else "The actual counts do not match the expected counts."`

Regarding the numerical variables: `r if (all_numeric(data_types)) "All numerical variables are correctly recognized as numerical." else "Some numerical variables are not recognized as numerical."`

#Q4
```{r}
num_observations <- nrow(data)

num_variables <- ncol(data)

num_missing_per_variable <- sapply(data, function(x) sum(is.na(x)))

num_obs_with_missing <- sum(apply(data, 1, function(x) any(is.na(x))))

summary_df <- data.frame(
  Variable = c("Number of Observations", "Number of Variables", "Observations with At Least One Missing Value"),
  MissingValues = c(num_observations, num_variables, num_obs_with_missing)
)

missing_values_df <- data.frame(
  Variable = names(num_missing_per_variable),
  MissingValues = num_missing_per_variable
)
missing_values_df <- missing_values_df[missing_values_df$MissingValues > 0, ]

final_table <- rbind(summary_df, missing_values_df)
```

```{r}
knitr::kable(final_table, caption = "Dataset Summary", format = "markdown")
```

#Q5
```{r}
data$luminosity <- as.numeric(data$luminosity)
data$`relative radius` <- as.numeric(data$`relative radius`)

data_clean <- data %>%
  filter(!is.na(luminosity) & !is.na(`relative radius`))

ggplot(data_clean, aes(x = `relative radius`, y = luminosity)) +
  geom_point() +
  labs(title = "Luminosity vs. Relative Radius",
       x = "Relative Radius",
       y = "Luminosity") +
  theme_minimal()

```
#Q6
```{r}
numerical_vars <- c("distance", "star temperature", "relative radius", "mass", "luminosity")

find_extremes <- function(var) {
  max_row <- data[which.max(data[[var]]), ]
  min_row <- data[which.min(data[[var]]), ]
  data.frame(
    Variable = var,
    "Max Star" = max_row$Star,
    "Max Value" = max_row[[var]],
    "Min Star" = min_row$Star,
    "Min Value" = min_row[[var]]
  )
}

extreme_values <- lapply(numerical_vars, find_extremes)

extreme_values_df <- do.call(rbind, extreme_values)
```


```{r}
knitr::kable(extreme_values_df, caption = "Most Extreme Stars for Each Numerical Variable", format = "markdown")

```

